Analysis of celiac disease

### 2018-06-07 3-simus.Rmd
some modifs by H
update

```{r, message=FALSE}
library(tidyverse)
library(glue)
NCORES <- 6
```

@@

```{r}
logit.CMSA <- function(G, pheno.all, covar.all, ind.train, ind.test,
                       method = "logit-simple", alpha = 0.5) {

@@

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               alpha = alpha, ncores = NCORES)

@@

```{r}
library(bigsnpr)
celiac2 <- snp_attach("backingfiles/celiacQC_sub1.rds")
CHR <- celiac2$map$chromosome
POS <- celiac2$map$physical.pos
(G <- celiac2$genotypes)
(G2 <- big_attach("backingfiles/celiacQC_sub1_tripled1.rds"))

@@

n <- nrow(G)
ind.HLA <- snp_indLRLDR(CHR, POS, subset(LD.wiki34, ID == "hild12"))
```

@@

```{r, eval=FALSE}
G3 <- big.matrix(nrow(G), ncol(G) + ncol(covar.all),
                 backingfile = "G-PC", 
                 backingpath = "backingfiles")
big_apply(G, function(X, ind) {
  G3[, ind] <- X[, ind]
  NULL
}, a.combine = 'c')
G3[, ncol(G) + cols_along(covar.all)] <- covar.all
```

```{r}
G3 <- attach.big.matrix("backingfiles/G-PC.desc")
```

@@

  # Get results from all methods
  res <- bind_rows(
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, 
               method = "logit-simple-enet-ridge", alpha = 0.001),
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, 
               method = "logit-simple",            alpha = 0.5),
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, 
               method = "logit-simple-lasso",      alpha = 1)
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

>>>>

```{r, message=FALSE}
library(tidyverse)
library(glue)
library(bigsnpr)
# https://stackoverflow.com/q/9486952/6103040
options(bigstatsr.cluster.type = "PSOCK")
```

@@

```{r}
logit.CMSA <- function(G, pheno.all, covar.all, ind.train, ind.test,
                       method = "logit-simple", alphas = 0.5) {

@@

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               alphas = alphas, ncores = NCORES)

@@

```{r}
celiac2 <- snp_attach("backingfiles/celiacQC_sub1.rds")
CHR <- celiac2$map$chromosome
POS <- celiac2$map$physical.pos
(G <- celiac2$genotypes)
(G2 <- big_attach("backingfiles/celiacQC_sub1_tripled1.rds"))

@@

n <- nrow(G)
ind.HLA <- snp_indLRLDR(CHR, POS, subset(LD.wiki34, ID == "hild12"))

NCORES <- nb_cores()
```

@@

```{r, eval=FALSE}
G3 <- bigmemory::big.matrix(nrow(G), ncol(G) + ncol(covar.all),
                            backingfile = "G-PC",                 
                            backingpath = "backingfiles")
big_apply(G, function(X, ind) {
  G3[, ind] <- X[, ind]
  NULL
}, a.combine = 'c')
G3[, ncol(G) + cols_along(covar.all)] <- covar.all
```

```{r}
G3 <- bigmemory::attach.big.matrix("backingfiles/G-PC.desc")
```

@@

  # Get results from all methods
  res <- bind_rows(
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, 
               method = "logit-simple-enet-ridge", alphas = 0.001),
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, 
               method = "logit-simple",            alphas = 0.5),
    logit.CMSA(G, pheno.all, covar.all, ind.train, ind.test, 
               method = "logit-simple-lasso",      alphas = 1)
  )
  params[["res"]] <- list(res)
  saveRDS(unnest(params, res, .drop = FALSE), file = res.file)
}
```

```{r, include=FALSE, error=TRUE}
library(gmailr)
send_message(
  mime(To = "florian.prive.21@gmail.com",
       From = "florian.prive.21@gmail.com",
       Subject = sprintf("Knitting of %s has finished",
                         knitr::current_input()))
)
```

### 2018-06-07 5-results-celiac.Rmd
update

```{r}
library(tidyverse)
myggplot <- function(..., coeff = 1) {
  bigstatsr:::MY_THEME(ggplot(...), coeff = coeff)
} 
```

@@

```{r, cache=FALSE}
# Read file, see "Preprocessing" notebook
library(bigsnpr)
celiac <- snp_attach("backingfiles/celiacQC.rds")
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos
NCORES <- nb_cores()
y <- celiac$fam$affection - 1
```

@@

```{r, cache=TRUE}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- apply(preds1, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds1.comb <- rowMeans(preds1)
AUC(preds1.comb, y[ind.test])
```

@@

```{r, cache=TRUE}
G2 <- big_attach("backingfiles/celiacQC_tripled1.rds")
system.time(
  logit2 <- big_spLogReg(X = G2, y01.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         ncores = NCORES)
)

@@

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               ncores = nb_cores())

@@

  # Get results from all methods
  res <- bind_rows(
    PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.CMSA(G2, pheno.all, covar.all, ind.train, ind.test, "logit-triple")
  )
  saveRDS(res, file = res.file)
}
```
```{r}
pAUC <- function(pred, target, p = 0.1) {
  val.min <- min(target)
  q <- quantile(pred[target == val.min], probs = 1 - p)
  ind <- (target != val.min) | (pred > q)
  tryCatch(bigstatsr::AUC(pred[ind], target[ind]) * p, 
           error = function(e) p^2 / 2)  # all same prediction
}

@@

list.files(path = "results6/", full.names = TRUE) %>%
  map_df(readRDS) %>%
  format_results() %>%
  group_by(method) %>%
  summarize_at(-1, function(x) {
    glue::glue("{signif(mean(x), 3)} ({signif(boot(x), 3)})")
  }) %>%
  print() %>%
  xtable::xtable()
```

>>>>

```{r}
library(tidyverse)
myggplot <- function(..., coeff = 1) {
  ggplot(...) + bigstatsr::theme_bigstatsr(size.rel = coeff)
} 
```

@@

```{r, cache=FALSE}
# Read file, see "Preprocessing" notebook
library(bigsnpr)
celiac <- snp_attach("backingfiles/celiacQC.rds")
G <- celiac$genotypes
CHR <- celiac$map$chromosome
POS <- celiac$map$physical.pos
NCORES <- nb_cores()
y <- celiac$fam$affection - 1

G2 <- big_attach("backingfiles/celiacQC_tripled1.rds")
```

@@

```{r, cache=TRUE}
# Train logit-simple
system.time(
  logit <- big_spLogReg(X = G, y01.train = y[ind.train], 
                        ind.train = ind.train, 
                        covar.train = obj.svd$u[ind.train, ],
                        ncores = NCORES, alpha = 0.5)
)
# Get K = 10 predictions/individuals
preds1 <- predict(logit, X = G, ind.row = ind.test, 
                  covar.row = obj.svd$u[ind.test, ])
(nb.pred1 <- sum(rowSums(sapply(logit, function(l) l$beta.X != 0)) > 0))
(aucs1 <- apply(preds1, 2, AUC, target = y[ind.test]))
# Combining all K predictions
preds1.comb <- rowMeans(preds1)
AUC(preds1.comb, y[ind.test])
```

@@

```{r, cache=TRUE}
system.time(
  logit2 <- big_spLogReg(X = G2, y01.train = y[ind.train], 
                         ind.train = ind.train, 
                         covar.train = obj.svd$u[ind.train, ],
                         ncores = NCORES, alphas = 0.5)
)

@@

    cmsa.logit <- big_spLogReg(X = G, y01.train = pheno.all[ind.train], 
                               ind.train = ind.train, 
                               covar.train = covar.all[ind.train, , drop = FALSE],
                               ncores = NCORES, alphas = 0.5)

@@

  # Get results from all methods
  res <- bind_rows(
    PRS(G, CHR, POS, pheno.all, covar.all, ind.train, ind.test),
    logit.CMSA(G,  pheno.all, covar.all, ind.train, ind.test, "logit-simple"),
    logit.CMSA(G2, pheno.all, covar.all, ind.train, ind.test, "logit-triple")
  )
  saveRDS(res, file = res.file)
}
```

```{r}
pAUC <- function(pred, target, p = 0.1) {
  val.min <- min(target)
  q <- quantile(pred[target == val.min], probs = 1 - p)
  ind <- (target != val.min) | (pred > q)
  tryCatch(bigstatsr::AUC(pred[ind], target[ind]) * p, 
           error = function(e) p^2 / 2)  # all same prediction
}

@@

list.files(path = "results6/", full.names = TRUE) %>%
  map_df(readRDS) %>%
  format_results() %>%
  group_by(method) %>%
  summarize_at(c("AUC", "pAUC", "nb.preds", "timing"), function(x) {
    glue::glue("{signif(mean(x), 3)} ({signif(boot(x), 3)})")
  }) %>%
  print() %>%
  xtable::xtable()
```
